% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\chapter{Generator-Pattern Repository}\label{chapter:generatorPatternRepository}
The following pages display the current status of the generator-pattern repository and its usage. It is optimized for the use with the regular expression framework used by the Java programming language.\\
When we applied the approach presented in this thesis we normalized the comments into \code{CloneChunk}s. We did so by using the procedure presented in section~\ref{section:convertToCloneChunks}.\\
The resulting \code{List} of \code{CloneChunk}s represents a \code{List} of single words, whereas all whitespaces and words without value have been removed. The resulting generator-pattern repository thus holds the generator-patterns without any formatting and programming language specific comment markers.

\section{Usage of the generator-pattern repository}
The easiest way to use the repository is by compiling the generator-patterns into regular expressions and normalizing the source code files when searching for the generator-patterns. This can be done in two steps:

\begin{enumerate}
	\item To use the generator-pattern \code{String}s one has to at first replace the placeholders \textit{<Version>, <Timestamp>, <Project>, <Class>} and \textit{<Link>}. The current observations have shown that a good replacement is the regex \textit{(.*)} that represents a sequence of any character with a variable length. Afterwards the resulting \code{String}s can be compiled into \code{Pattern}s and used in the regex framework.
	\item Due to the fact that the formatting of the generator-patterns has been removed to have them as universally as possible the source code files have to be normalized. \\
	At first one has to split the source code files at every linebreak and whitespace. Afterwards, every word is checked if it is valuable (see section~\ref{section:convertToCloneChunks}). The remaining words get merged into one \code{String}, whereas the words are delimited by a single whitespace. An exemplary Java implementation is displayed in listing~\ref{lst:normalizeFunction}.\\
	On the resulting  \code{String} a regex matching can be performed with the generator-patterns.
\end{enumerate}
\input{listings/normalizeFunction.tex}

\cleardoublepage{}
\section{Qualitas Corpus}
The following tables picture the set of generator-patterns that was found using the algorithm on the Qualitas Corpus.
\input{tables/generatorPatternRepositoryQC.tex}
\input{tables/generatorPatternRepositoryQC_01.tex}
\input{tables/generatorPatternRepositoryQC_02.tex}

\cleardoublepage{}
\section{Random projects}
The following tables picture the set of generator-patterns that was found using the algorithm on the random projects.
\input{tables/generatorPatternRepositoryGit.tex}
\input{tables/generatorPatternRepositoryGit_01.tex}